---
title: "Advent of Code 2018"
author: Anil Goyal
output:
  md_document:
    variant: markdown_github
---

```{r}
library(tidyverse)
library(unglue)
library(lubridate)
library(data.table)
```

### Day 1
### [--- Day 1: Chronal Calibration ---](https://adventofcode.com/2018/day/1)

#### Part-1
```{r}
input <- read_lines('input1.txt')

answer1_1 <- function(input){
  input %>% 
    as.integer() %>% 
    sum()
}

answer1_1(input)
```

#### Part-2
```{r}
answer1_2 <- function(input){
  freq <- input %>% as.integer()
  # initiate while loop
  while(!any(duplicated(cumsum(freq)))){
    freq <- append(freq, freq)
  }
  # final answer
  cumsum(freq)[which.max(duplicated(cumsum(freq)))]
}

answer1_2(input)
```


### Day-2
### [--- Day 2: Inventory Management System ---](https://adventofcode.com/2018/day/2)

#### Part-1
```{r}
input <- read_lines('input2.txt')

answer2_1 <- function(input){
  twos <- input %>% 
    str_split('') %>% 
    map_lgl(~ 2 %in% table(.x)) %>% 
    sum()
  
  threes <- input %>% 
    str_split('') %>% 
    map_lgl(~ 3 %in% table(.x)) %>% 
    sum() 
  
  twos * threes
}

answer2_1(input)
```

#### Part-2
```{r}
answer2_2 <- function(input){
  # modify input
  input <- input %>% str_split('')
  # initiate while loop to find i
  i <- 1
    while(!any(map_int(input[-i], ~sum(.x == input[[i]])) == 25)){
    i <- i+1
  }
  my_char <- input[[i]]
  my_char2 <- input[-i][[which.max(map_int(input[-i], ~ sum(.x == input[[i]])))]]
  # answer
  paste0(my_char[my_char == my_char2], collapse = '')
}

answer2_2(input)
```


### Day-3
### [--- Day 3: No Matter How You Slice It ---](https://adventofcode.com/2018/day/3)

#### Part-1
```{r}
input <- read_lines('input3.txt')

answer3_1 <- function(input){
  # patterns for ungluing details
  patterns <- '#{carpet_id=\\d+} @ {mar_w=\\d+},{mar_h=\\d+}: {carpet_w=\\d+}x{carpet_h=\\d+}'
  # database
  dat <- unglue_data(input, patterns = patterns, convert = TRUE)
  # define helper function
  carpet <- function(mw, mh, cw, ch){
    s_p <- complex(real = mw, imaginary = mh)
    s_ps <- seq(from=s_p, by = 1+0i, length.out=cw)
    map(s_ps, ~ seq(.x, by = 0+1i, length.out=ch)) %>% 
      unlist
  }
  
  # all carpets
  all_carps <- map(seq_len(nrow(dat)), ~carpet(dat$mar_w[.x], dat$mar_h[.x], dat$carpet_w[.x], dat$carpet_h[.x]))
  # answer
  all_carps %>% unlist %>% 
    table %>% 
    {which(. > 1)} %>% 
    length
}

answer3_1(input)
```

#### Part-2
```{r}
answer3_2 <- function(input){
  # patterns for ungluing details
  patterns <- '#{carpet_id=\\d+} @ {mar_w=\\d+},{mar_h=\\d+}: {carpet_w=\\d+}x{carpet_h=\\d+}'
  # database
  dat <- unglue_data(input, patterns = patterns, convert = TRUE)
  # define helper function
  carpet <- function(mw, mh, cw, ch){
    s_p <- complex(real = mw, imaginary = mh)
    s_ps <- seq(from=s_p, by = 1+0i, length.out=cw)
    map(s_ps, ~ seq(.x, by = 0+1i, length.out=ch)) %>% 
      unlist
  }
  
  # all carpets
  all_carps <- map(seq_len(nrow(dat)), ~carpet(dat$mar_w[.x], dat$mar_h[.x], dat$carpet_w[.x], dat$carpet_h[.x]))
  # Overlapping area
  overlap <- all_carps %>% unlist %>% 
    table %>% 
    {which(. > 1)} %>% 
    names() %>% 
    as.complex()
  
  # answer
  map_lgl(all_carps, ~ !any(.x %in% overlap)) %>% 
    which.max()
  
}

answer3_2(input)
```


### Day-4
### [--- Day 4: Repose Record --](https://adventofcode.com/2018/day/4)

#### Part-1
**Explanation** A solution using `dplyr` and `tidyr` strategy.  `data.table::rleid()` has also been used.
```{r}
input <- read_lines('input4.txt')

answer4_1 <- function(input){
  # patterns
  patterns <- c('[{time}] Guard #{guard_no=\\d+} {pat}',
                '[{time}] {pat}')
  # unglue data
  dat <- unglue::unglue_data(input, patterns = patterns) %>% 
    mutate(time = ymd_hm(time))
  
  # modify data to intermediate usable format
  dat2 <- dat %>% 
    arrange(time) %>% 
    fill(guard_no, .direction= 'down') %>% 
    mutate(duty = rleid(guard_no)) %>% 
    group_by(duty, guard_no) %>% 
    complete(time = seq.POSIXt(from = min(time), to = max(time), by = 'min')) %>% 
    fill(pat, .direction = 'down') %>% 
    ungroup %>% 
    filter(pat == 'falls asleep') %>% 
    mutate(minute = minute(time))  
  
  # guard number (strategy-1)
  guard <- dat2 %>% 
    group_by(guard_no) %>% 
    summarise(asleep = n()) %>% 
    filter(asleep == max(asleep)) %>% 
    pull(guard_no) %>% 
    as.numeric()
  # desired minute 
  minute <- dat2 %>% 
    filter(guard_no == guard) %>% 
    group_by(minute) %>% 
    summarise(times = n()) %>% 
    filter(times == max(times)) %>% 
    pull(minute)
  # answer
  guard * minute
}
answer4_1(input)
```

#### Part-2
```{r}
answer4_2 <- function(input){
  # patterns
  patterns <- c('[{time}] Guard #{guard_no=\\d+} {pat}',
                '[{time}] {pat}')
  # unglue data
  dat <- unglue::unglue_data(input, patterns = patterns) %>% 
    mutate(time = ymd_hm(time))
  
  # modify data to intermediate usable format
  dat2 <- dat %>% 
    arrange(time) %>% 
    fill(guard_no, .direction= 'down') %>% 
    mutate(duty = rleid(guard_no)) %>% 
    group_by(duty, guard_no) %>% 
    complete(time = seq.POSIXt(from = min(time), to = max(time), by = 'min')) %>% 
    fill(pat, .direction = 'down') %>% 
    ungroup %>% 
    filter(pat == 'falls asleep') %>% 
    mutate(minute = minute(time)) 
  
  # strategy-2 is easier
  dat2 %>% 
  group_by(guard_no, minute) %>% 
  summarise(times_asleep = n(), .groups = 'drop') %>% 
  filter(times_asleep == max(times_asleep)) %>% 
  {as.numeric(.$guard_no)*.$minute}
}

answer4_2(input)
```


### Day-5
### [--- Day 5: Alchemical Reduction ---](https://adventofcode.com/2018/day/5)

#### Part-1
```{r}
input <- read_lines('input5.txt')

answer5_1 <- function(input){
  # starting polymer
  polymer <- input
  # reactive polymers
  reactive_polymers <- c(paste0(letters, LETTERS), paste0(LETTERS, letters))
  # while loop to reduce polymer to unreactive polymer
  while(any(map_lgl(reactive_polymers , ~str_detect(polymer, .x)))){
    polymer <- reduce(reactive_polymers, 
                      .init = polymer,
                      ~str_remove_all(.x, .y))
  }
  # length of unreactive polymer
  nchar(polymer)
}

answer5_1(input)
```

#### Part-2
```{r}
answer5_2 <- function(input){
  # get polymer
  polymer <- input
  # reactive polymers list
  reactive_polymers <- c(paste0(letters, LETTERS), paste0(LETTERS, letters))
  # new reduction list
  reduction_list <- split(c(LETTERS, letters), rep(1:26, 2))
  # helper function this time
  my_fun <- function(polymer, reactive_polymers = reactive_polymers){
    while(any(map_lgl(reactive_polymers , ~str_detect(polymer, .x)))){
      polymer <- reduce(reactive_polymers, 
                        .init = polymer,
                        ~str_remove_all(.x, .y))
    }
    nchar(polymer)
  }
  # answer
  map(reduction_list, function(.a) reduce(.a, .init = polymer, ~str_remove_all(.x, .y))) %>% 
    map_int(my_fun, reactive_polymers) %>% 
    min()
}

answer5_2(input)
```


### Day-6
### [--- Day 6: Chronal Coordinates ---](https://adventofcode.com/2018/day/6)

#### Part-1
```{r}
input <- read_lines('input6.txt')

answer6_1 <- function(input){
  
  patterns <- '{x}, {y}'
  
  dat <- unglue_data(input, patterns, convert = TRUE)
  
  dat <- dat %>% 
    mutate(z= complex(real = x, imaginary = y))
  
  man_dist <- function(x, y){
    diff(sort(c(Re(x), Re(y)))) + 
      diff(sort(c(Im(x), Im(y))))
  }
  
  borders <- function(z){
    min_x <- min(Re(z))
    max_x <- max(Re(z))
    min_y <- min(Im(z))
    max_y <- max(Im(z))
    f <- seq(complex(real = min_x, imaginary = min_y), by = 1, length.out = max_x - min_x + 1)
    s <- seq(complex(real = min_x, imaginary = min_y), by = 0+1i, length.out = max_y - min_y + 1)
    t <- seq(tail(f, 1), by = 0+1i, length.out = max_y - min_y + 1)
    l <- seq(tail(s,1), by = 1, length.out = max_x - min_x + 1)
    unique(c(f, s, t, l))
  }
  
  finite_points <- function(z){
    min_x <- min(Re(z))
    max_x <- max(Re(z))
    min_y <- min(Im(z))
    max_y <- max(Im(z))
    
    expand_grid(x=seq(min_x+1, max_x-1, by=1), 
                y=seq(min_y+1, max_y-1, by=1)) %>% 
      transmute(z = complex(real = x, imaginary = y)) %>% 
      {.$z}
  }
  
  my_fun <- function(z){
    dist_vec <- map_dbl(dat$z, ~man_dist(.x, z))
    if(length(which(dist_vec == min(dist_vec))) == 1){
      dat$z[which(dist_vec == min(dist_vec))]
    } else {
      NA
    }
  }
  res <- map(finite_points(dat$z), 
             my_fun ) %>% unlist
  bor <- map(borders(dat$z), my_fun) %>% unlist
  # answer
  table(res)[!names(table(res)) %in% names(table(bor))] %>% max
}

answer6_1(input)
```

#### Part-2
```{r}
answer6_2 <- function(input){
  patterns <- '{x}, {y}'
  
  dat <- unglue_data(input, patterns, convert = TRUE)
  
  dat <- dat %>% 
    mutate(z= complex(real = x, imaginary = y))
  
  man_dist <- function(x, y){
    diff(sort(c(Re(x), Re(y)))) + 
      diff(sort(c(Im(x), Im(y))))
  }
  
  finite_points <- function(z){
    min_x <- min(Re(z))
    max_x <- max(Re(z))
    min_y <- min(Im(z))
    max_y <- max(Im(z))
    
    expand_grid(x=seq(min_x+1, max_x-1, by=1), 
                y=seq(min_y+1, max_y-1, by=1)) %>% 
      transmute(z = complex(real = x, imaginary = y)) %>% 
      {.$z}
  }
  
  my_fun <- function(z){
    sum(map_dbl(dat$z, ~man_dist(.x, z))) < 10000
  }
  map_lgl(finite_points(dat$z), 
          my_fun ) %>% sum
  
}

answer6_2(input)
```


### Day-7
### [--- Day 7: The Sum of Its Parts ---](https://adventofcode.com/2018/day/7)

#### Part-1
```{r}
input <- read_lines('input7.txt')

answer7_1 <- function(input){
  patterns <- 'Step {preq} must be finished before step {post} can begin.'
  
  dat <- unglue_data(input, patterns)
  
  all_tasks <- union(dat$preq, dat$post)
  finished <- c()
  for(i in seq_along(all_tasks)){
    available <- setdiff(all_tasks, unique(dat$post[!dat$preq %in% finished]))
    finished <- append(finished, sort(setdiff(available, finished))[1])
  }
  #answer
  paste0(finished, collapse = '')
}

answer7_1(input)
```

#### Part-2
```{r}

```

### Day-8
### [](https://adventofcode.com/2018/day/8)

#### Part-1
```{r}

```

#### Part-2
```{r}

```

### Day-9
### [](https://adventofcode.com/2018/day/9)

#### Part-1
```{r}

```

#### Part-2
```{r}

```

### Day-10
### [](https://adventofcode.com/2018/day/10)

#### Part-1
```{r}

```

#### Part-2
```{r}

```

### Day-11
### [--- Day 11: Chronal Charge ---](https://adventofcode.com/2018/day/11)

#### Part-1
**Explanation:**  Solved this with use of complex numbers which can be thought of two dimensional numbers every time.
```{r}
input <- 9445
answer11_1 <- function(input=9445){
  # helper function
  grid33 <- function(x){
    e <- expand.grid(0:2,0:2)
    x + complex(real=e$Var1, imaginary = e$Var2)
  }
  # Complete input grid (Leave last two)
  input_grid <- expand.grid(1L:298L, 1L:298L) %>% 
    transmute(x = complex(real=Var1, imaginary = Var2)) %>% 
    pull(x)
  # power of one cell
  power_1 <- function(x){
    rack <- Re(x) + 10
    (((rack * Im(x) +input)*rack) %/% 100) %% 10
  }
  #Vectorise above helper function
  power_g <- Vectorize(power_1, vectorize.args = 'x')
  
  map_int(input_grid, ~ grid33(.x) %>% 
            power_g() %>% 
            sum() %>% 
            as.integer) %>% 
    which.max() %>% 
    {input_grid[.]} %>% 
    {paste(Re(.), Im(.), sep = ',')}
  
}

answer11_1()
```

#### Part-2
```{r}

```

### Day-12
### [](https://adventofcode.com/2018/day/12)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


