---
title: "Advent of Code 2016"
author: Anil Goyal
output:
  md_document:
    variant: markdown_github
---

```{r}
library(tidyverse)
library(unglue)
```

### Day 1
### [--- Day 1: No Time for a Taxicab ---](https://adventofcode.com/2016/day/1)

#### Part-1
**Explanation** - I am doing it with the use of complex numbers.  Complex numbers can be interpreted as two dimensional numbers.  So I will use these as location of Santa. Needless to say real part will denote $x$ coordinate and imaginary part will denote $y$ coordinate.  Since initially he is facing north, `.init` in `purrr::accumulate` has been used as $0+1i$.
```{r}
input <- readLines('input1.txt')

answer1_1 <- function(input){
  # parse input in a more usable way
  dat <- str_split(input, ', ') %>% unlist
  
  # Define a helper function for purrr::accumulate
  my_fun <- function(.x, .y){
    steps <- as.numeric(gsub('R|L', '', .y))
    .x * if(startsWith(.y, 'R')){
      0-1i 
    } else {
      0+1i 
    }
  }
  # x vector will have directions only
  x <- accumulate(dat, .init = 0+1i, my_fun)
  
  # y vector will have distance travelled
  y <- c(0, as.numeric(gsub('R|L', '', dat)))
  
  # z is an intermediate value
  z <- reduce(x*y, `+`)
  
  # answer
  abs(Re(z))+abs(Im(z))
}
answer1_1(input)
```

#### Part-2
**Explanation** - Here we need complete path travelled.  We have directions to travel, initial point and final points.  Using these values through `seq` of baser, I have build complete locations travelled using `purrr::map` and `unlist`.  Last part is then easy.
```{r}
answer1_2 <- function(input){
  # parse input in a more usable way
  dat <- str_split(input, ', ') %>% unlist
  
  # Define a helper function for purrr::accumulate
  my_fun <- function(.x, .y){
    steps <- as.numeric(gsub('R|L', '', .y))
    .x * if(startsWith(.y, 'R')){
      0-1i 
    } else {
      0+1i 
    }
  }
  # x vector will have directions only
  x <- accumulate(dat, .init = 0+1i, my_fun)
  
  # y vector will have distance travelled
  y <- c(0, as.numeric(gsub('R|L', '', dat)))
  
  # locations travelled
  z <- cumsum(x*y)
  
  #complete itinerary
  complete_map <- unlist(map(seq_along(x), ~seq(to=z[.x], by=x[.x], length.out=y[.x])))
  
  # office location
  office <- complete_map[which.max(duplicated(complete_map))]
  
  # answer
  abs(Re(office))+abs(Im(office))
  
}
answer1_2(input)
```

### Day-2
### [--- Day 2: Bathroom Security ---](https://adventofcode.com/2016/day/2)