---
title: "Advent of Code 2015"
author: Anil Goyal
output:
  md_document:
    variant: markdown_github
---

```{r}
library(tidyverse)
library(digest)
library(unglue)
```

### Day 1
### [Not Quite Lisp](https://adventofcode.com/2015/day/1)

#### Part-1

```{r}
input <- readLines('input1.txt')
# part-1
answer1_1 <- function(input){
  input_vec <- input %>% str_split('') %>% unlist
  floors <- c(-1, 1)[1+(input_vec == '(')]
  sum(floors)
}

answer1_1(input)
```

#### part-2
```{r}
answer1_2 <- function(input){
  input_vec <- input %>% str_split('') %>% unlist
  floors <- c(-1, 1)[1+(input_vec == '(')]
  # answer
  which.max(cumsum(floors)<0)
}

answer1_2(input)
```


### Day2
### [I Was Told There Would Be No Math](https://adventofcode.com/2015/day/2)

#### Part-1

```{r}
input <- readLines('input2.txt')

#part-1
answer2_1 <- function(input){
  input_list <- input %>% 
    str_split('x') %>% 
    map(as.integer)
  
  
  l <- map_int(input_list, ~ sort(.x)[1])
  b <- map_int(input_list, ~ sort(.x)[2])
  h <- map_int(input_list, ~ sort(.x)[3])
  #answer
  sum(3*l*b + 2*l*h + 2*b*h)
}

answer2_1(input)
```


#### Part-2

```{r}
answer2_2 <- function(input){
  input_list <- input %>% 
    str_split('x') %>% 
    map(as.integer)
  
  
  l <- map_int(input_list, ~ sort(.x)[1])
  b <- map_int(input_list, ~ sort(.x)[2])
  h <- map_int(input_list, ~ sort(.x)[3])
  #answer
  sum(2*(l+b)+l*b*h)
}

answer2_2(input)
```



### Day 3 
### [Perfectly Spherical Houses in a Vacuum](https://adventofcode.com/2015/day/3) 

#### Part-1
```{r}
input <- readLines('input3.txt')
## part-1
answer3_1 <- function(input){
  input_vec <- input %>% 
    str_split('') %>% 
    unlist
  
  input_vec <- ifelse(input_vec== '>', 1, 
                      ifelse(input_vec == '<', -1, 
                             ifelse(input_vec == '^', 1i, -1i)))
  
  input_vec <- c(0+0i, input_vec)
  # answer
  length(unique(cumsum(input_vec)))
}

answer3_1(input)
```

#### Part-2

```{r}
answer3_2 <- function(input){
  input_vec <- input %>% 
    str_split('') %>% 
    unlist
  
  input_vec <- ifelse(input_vec== '>', 1, 
                      ifelse(input_vec == '<', -1, 
                             ifelse(input_vec == '^', 1i, -1i)))
  #answer
  split(input_vec, seq_along(input_vec) %% 2) %>% 
    map(~ c(0+0i, .x) %>% 
          cumsum()) %>% 
    reduce(union) %>% 
    length
}

answer3_2(input)
```


### Day-4 
### [The Ideal Stocking Stuffer](https://adventofcode.com/2015/day/4)

#### Part-1
```{r}
answer4_1 <- function(first_part){
  second_part <- 1
  while(!startsWith(digest(paste0(first_part, second_part), algo = 'md5', serialize = FALSE), '00000')){
    second_part <- second_part + 1
  }
  second_part
}

answer4_1('ckczppom')

```

#### Part-2

```{r}
answer4_2 <- function(first_part){
  second_part <- 1
  while(!startsWith(digest(paste0(first_part, second_part), algo = 'md5', serialize = FALSE), '000000')){
    second_part <- second_part + 1
  }
  second_part
}

answer4_2('ckczppom')
```

### Day 5
### [Doesn't He Have Intern-Elves For This?](https://adventofcode.com/2015/day/5)

#### Part-1

```{r}
input <- readLines('input5.txt')
answer5_1 <- function(input){
  # vowels
  rule1_st <- c('a', 'e', 'i', 'o', 'u')
  # check for at least 3 vowels
  rule1 <- map_lgl(str_split(input, '') , ~ sum(.x %in% rule1_st) >= 3)
  # check for repeated alphabets
  rule2 <- map_lgl(str_split(input, ''), function(.a)any(map_lgl(seq(length(.a)-1), ~.a[.x]==.a[.x+1])))
  # rule 3 exclusion strings
  rule3_st <- c('ab', 'cd', 'pq', 'xy')
  # check for exclusions
  rule3 <- !map_lgl(input, function(.a)any(map_lgl(rule3_st, ~str_detect(.a, .x))))
  # answer (nice strings)
  sum(rule1 & rule2 & rule3)
}

answer5_1(input)
```

#### Part-2

```{r}

```


### Day 6
### [Day 6: Probably a Fire Hazard](https://adventofcode.com/2015/day/6)

#### Part-1
```{r}
input <- readLines('input6.txt')

answer6_1 <- function(input){
  
  # first define a helper function
  my_fun <- function(.x, .y){
    if (.y$ins == 'turn on'){
      .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] <- TRUE
      .x
    } else if (.y$ins == 'turn off'){
      .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] <- FALSE
      .x
    } else {
      .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] <- !(.x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)])
      .x
    }
  }
  
  # read the input properly
  patterns <- '{ins} {x1=\\d+},{y1=\\d+} through {x2=\\d+},{y2=\\d+}'
  input_inst <- unglue::unglue(input, patterns, convert = TRUE)
  
  #prepare initial configuration of lights
  mat <- matrix(FALSE, nrow = 1000, ncol=1000)
  
  #iteration
  output <- reduce(input_inst, .init = mat,
       my_fun)
  
  # answer
  sum(output)
  
}

answer6_1(input)

```

#### Part-2

It has good usage of r's `pmax` function
```{r}
input <- readLines('input6.txt')

answer6_2 <- function(input){
  
  # first define a helper function
  my_fun <- function(.x, .y){
    if (.y$ins == 'turn on'){
      .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] <- .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] + 1L
      .x
    } else if (.y$ins == 'turn off'){
      .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] <- pmax(0, .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] -1L)
      .x
    } else {
      .x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)] <- (.x[seq(.y$x1 +1, .y$x2 +1, 1), seq(.y$y1 +1, .y$y2 +1, 1)]) + 2L
      .x
    }
  }
  
  # read the input properly
  patterns <- '{ins} {x1=\\d+},{y1=\\d+} through {x2=\\d+},{y2=\\d+}'
  input_inst <- unglue::unglue(input, patterns, convert = TRUE)
  
  #prepare initial configuration of lights
  mat <- matrix(FALSE, nrow = 1000, ncol=1000)
  
  #iteration
  output <- reduce(input_inst, .init = mat,
       my_fun)
  
  # answer
  sum(output)
}

answer6_2(input)

```


### Day-7
### [Day 7: Some Assembly Required](https://adventofcode.com/2015/day/7)

#### Part-1

```{r}

```


### Day-8


### Day-9


### Day-10

### [--- Day 10: Elves Look, Elves Say ---](https://adventofcode.com/2015/day/10)

#### Part-1
```{r}
answer10_1 <- function(input){
  nchar(reduce(1:40, .init = input, ~ .x %>% str_split('') %>%                # used purrr::reduce instead of for loop just for personal liking  
  unlist %>% 
  rle %>%                                                                     # used base r's rle function
  {c(rbind(as.character(.$lengths), .$values))} %>%                           # interleaved two vectors by c and rbind
    paste0(collapse = '')))                                                   # pasted them back
}

answer10_1('1113122113')
```

#### Part-II
```{r}
answer10_2 <- function(input){
  nchar(reduce(1:50, .init = input, ~ .x %>% str_split('') %>% 
  unlist %>% 
  rle %>% 
  {c(rbind(as.character(.$lengths), .$values))} %>% 
    paste0(collapse = '')))
}

answer10_2('1113122113')
```



### Day-11
### [--- Day 11: Corporate Policy ---](https://adventofcode.com/2015/day/11)