---
title: "Advent of Code 2020"
author: Anil Goyal
output:
  md_document:
    variant: markdown_github
---

```{r}
library(tidyverse)
library(unglue)

```


### Day 1
### [--- Day 1: Report Repair ---](https://adventofcode.com/2020/day/1)

#### Part-1
```{r}
input <- read_lines('input1.txt') %>% as.integer()

answer1_1 <- function(input){
  
  combn(input, 2) %>% 
    {.[,which(colSums(.) == 2020)]} %>% 
    prod()
}
answer1_1(input)
```

#### Part-2
```{r}
answer1_2 <- function(input){
  
  combn(input, 3) %>% 
    {.[,which(colSums(.) == 2020)]} %>% 
    prod()
}
answer1_2(input)
```


### Day 2
### [--- Day 2: Password Philosophy ---](https://adventofcode.com/2020/day/2)

#### Part-1
```{r}
input <- read_lines('input2.txt')

answer2_1 <- function(input){
  patterns <- '{min=\\d+}-{max=\\d+} {char=\\w}: {word}'
  
  dat <- unglue::unglue_data(input, patterns, convert = T)
  
  dat %>% 
    summarise(valid_pw = sum(str_count(word, char) >= min & str_count(word, char) <= max)) %>% 
    pull(valid_pw)
}
answer2_1(input)
```

#### Part-2
```{r}
answer2_2 <- function(input){
  patterns <- '{min=\\d+}-{max=\\d+} {char=\\w}: {word}'
  
  dat <- unglue::unglue_data(input, patterns, convert = T)
  
  dat %>% 
    summarise(valid_pw = sum(xor(substr(word, min, min)==char, 
                                 substr(word, max, max) == char))) %>% 
    pull(valid_pw)
  
}
answer2_2(input)
```


### Day 3
### [--- Day 3: Toboggan Trajectory ---](https://adventofcode.com/2020/day/3)

#### Part-1
```{r}
input <- readLines('input3.txt')
answer3_1 <- function(input){
    mat.in <- input %>% 
    str_split('') %>% 
    unlist %>% 
    matrix(nrow = length(input), byrow = TRUE)
  
  sum(map_chr(seq(nrow(mat.in)), ~mat.in[.x, (((.x -1)*3 ) %% ncol(mat.in)) +1 ]) == '#')
}

answer3_1(input)
```

#### Part-2
```{r}
answer3_2 <- function(input){
  mat.in <- input %>% 
    str_split('') %>% 
    unlist %>% 
    matrix(nrow = length(input), byrow = TRUE)
  #calculating five slope tree maps
  slope1 <- sum(map_chr(seq(nrow(mat.in)), ~mat.in[.x, (((.x -1)*1 ) %% ncol(mat.in)) +1 ]) == '#')
  slope2 <- sum(map_chr(seq(nrow(mat.in)), ~mat.in[.x, (((.x -1)*3 ) %% ncol(mat.in)) +1 ]) == '#')
  slope3 <- sum(map_chr(seq(nrow(mat.in)), ~mat.in[.x, (((.x -1)*5 ) %% ncol(mat.in)) +1 ]) == '#')
  slope4 <- sum(map_chr(seq(nrow(mat.in)), ~mat.in[.x, (((.x -1)*7 ) %% ncol(mat.in)) +1 ]) == '#')  
  slope5 <- sum(map_chr(seq(1+nrow(mat.in)/2), ~mat.in[2*.x -1, ((.x -1) %% ncol(mat.in)) +1 ]) == '#')
  #final result
  prod(map_int(1:5, ~ get(paste0('slope', .x))))
  
}

answer3_2(input)
```


### Day 4
### [](https://adventofcode.com/2020/day/4)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


### Day 5
### [](https://adventofcode.com/2020/day/5)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


### Day 6
### [--- Day 6: Custom Customs ---](https://adventofcode.com/2020/day/6)

#### Part-1
```{r}

input <- readLines('input6.txt')
answer6_1 <- function(x){
  
  i1 <- !nzchar(x)
  inp_l <- unname(split(x[!i1], cumsum(i1)[!i1]))
  
  sum(map_int(inp_l, ~reduce(str_split(.x, ''), union) %>% length))
}

answer6_1(input)
```

#### Part-2
```{r}
answer6_2 <- function(x){
  
  i1 <- !nzchar(x)
  inp_l <- unname(split(x[!i1], cumsum(i1)[!i1]))
  
  sum(map_int(inp_l, ~reduce(str_split(.x, ''), intersect) %>% length))
}

answer6_2(input)
```


### Day 7
### [](https://adventofcode.com/2020/day/7)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


### Day 8
### [--- Day 8: Handheld Halting ---](https://adventofcode.com/2020/day/8)

#### Part-1
```{r}
input <- read_lines('input8.txt')

answer8_1 <- function(input){
  patterns <- '{op} {amt}'
  dat <- unglue::unglue_data(input, patterns = patterns, convert = TRUE)
  # initial values for iteration
  acc <- 0
  visits <- c()
  cur_state <- 1
  # While loop
  while(! cur_state %in% visits){
    if(dat$op[cur_state] == 'acc'){
      visits <- append(visits, cur_state)
      acc <- acc + dat$amt[cur_state]
      cur_state <- cur_state +1
    } else if (dat$op[cur_state] == 'jmp'){
      visits <- append(visits, cur_state)
      cur_state <- cur_state + dat$amt[cur_state]
    } else {
      visits <- append(visits, cur_state)
      cur_state <- cur_state +1
    }
  }
  # answer
  acc
}

answer8_1(input)
```

#### Part-2
```{r}

```


### Day 9
### [--- Day 9: Encoding Error ---](https://adventofcode.com/2020/day/9)

#### Part-1
```{r}
input <- read_lines('input9.txt') %>% as.numeric()

answer9_1 <- function(input, preamble=25){
  # my preamble
  n <- preamble
  # seq
  ss <- seq(n+1,length(input), by = 1)
  # helper function
  my_fun <- function(.x){
    prev <- seq(.x-n, .x-1, by=1)
    
    min_r <- input[prev] %>% 
      sort() %>% 
      head(2) %>% 
      sum()
    
    max_r <- input[prev] %>% 
      sort() %>% 
      tail(2) %>% 
      sum()
    
    !between(input[.x], min_r, max_r)
  }
  # final answer
  input[ss[map_lgl(ss, my_fun)]]
}

answer9_1(input, 25)
```

#### Part-2
```{r}

```


### Day 10
### [--- Day 10: Adapter Array ---](https://adventofcode.com/2020/day/10)

#### Part-1
```{r}
input <- read_lines('input10.txt')
answer10_1 <- function(input){
  input %>% 
    as.integer() %>% 
    {c(0L, . , max(.)+3L)} %>% 
    sort() %>% 
    diff %>% 
    table %>% 
    {.['1']*.['3']} %>% 
    unname()
}

answer10_1(input)
```

#### Part-2
Explanation:  By a simple hit and trial we may know that there are 

 - no choices whenever difference is 3
 - 2 choices when there are 2 consecutive 1s
 - 4 choices when there are 3 consecutive 1s
 - 7 choices when there are 4 consecutive 1s
```{r}
answer10_2 <- function(input){
  input %>% 
  as.integer() %>% 
  {c(0L, . , max(.)+3L)} %>% 
  sort() %>% 
  diff %>% 
  rle() %>% 
  {.$lengths[.$values == 1]} %>% 
  {ifelse(.==4, 7, ifelse(. == 3, 4, .))} %>% 
  prod %>% 
  {print(., digits = 13)}
}

answer10_2(input)
```


### Day 11
### [](https://adventofcode.com/2020/day/11)

#### Part-1
```{r}

```

#### Part-2
```{r}

```



### Day 12
### [](https://adventofcode.com/2020/day/12)

#### Part-1
```{r}

```

#### Part-2
```{r}

```



### Day 13
### [](https://adventofcode.com/2020/day/13)

#### Part-1
```{r}

```

#### Part-2
```{r}

```



### Day 14
### [](https://adventofcode.com/2020/day/14)

#### Part-1
```{r}

```

#### Part-2
```{r}

```



### Day 15
### [--- Day 15: Rambunctious Recitation ---](https://adventofcode.com/2020/day/15)

**Explanation** This is [Van Eck sequence](https://oeis.org/A181391).

#### Part-1
```{r}
answer15_1 <- function(my_seq, number){
  while(length(my_seq)<number){
    if(!my_seq[length(my_seq)] %in% my_seq[-length(my_seq)]){
      my_seq <- append(my_seq, 0L)
    } else {
      my_seq <- append(my_seq, length(my_seq) - max(which(my_seq[-length(my_seq)] == my_seq[length(my_seq)])))
    }
  }
  
  my_seq[length(my_seq)]
}

answer15_1(c(15L, 5L, 1L, 4L, 7L, 0L), number = 2020L)
```

**Note:** Though part-1 of the answer/code above may also work but it will be optimised to have only last indices of unique number in each iteration.  So, a different function has been used, which aid to generate only the current/last number or alternatively the last index of any unique number in the sequence.

Hints have been taken from [https://github.com/fdlk/advent-2020/blob/master/day15.md](https://github.com/fdlk/advent-2020/blob/master/day15.md)

#### Part-2
```{r}
input <- c(15L, 5L, 1L, 4L, 7L, 0L)

answer15_2 <- function(input){
  game <- rep(0, 30000000)
  for (i in 1:(length(input) - 1)) {
    number <- input[[i]]
    game[number + 1] <- i
  }
  
  
  last_number <- last(input)
  
  
  for (i in (length(input) + 1):30000000) {
    occurred_last <- game[[last_number + 1]]
    game[[last_number + 1]] <- i - 1
    last_number <- if (occurred_last == 0) {
      0
    } else {
      i - 1 - occurred_last
    }
  }
  
  
  last_number
  
}

answer15_2(input)
```



### Day 16
### [](https://adventofcode.com/2020/day/16)

#### Part-1
```{r}

```

#### Part-2
```{r}

```



### Day 17
### [](https://adventofcode.com/2020/day/17)

#### Part-1
```{r}

```

#### Part-2
```{r}

```



### Day 18
### [--- Day 18: Operation Order ---](https://adventofcode.com/2020/day/18)

#### Part-1
```{r}
input <- read_lines('input18.txt')

answer18_1 <- function(input){
  # define helper functions
  `%a%` <- function(x,y) x+y
  `%s%` <- function(x,y) x-y
  `%m%` <- function(x,y) x*y
  `%d%` <- function(x,y) x/y
  
  # modify input
  input <- reduce2(list('\\+', '\\-', '\\*', '\\/'), list('%a%', '%s%', '%m%', '%d%'),
                   .init = input,
                   function(.x, .y, .z) map(.x, function(.a) str_replace_all(.a, .y, .z )))
  
  # final answer
  map_dbl(input, ~ eval(parse(text = .x))) %>% 
    sum()
  
}

answer18_1(input)
```

#### Part-2
```{r}
answer18_2 <- function(input){
  # helper function
  `/` <- function(x,y) x+y    
  `-` <- function(x,y) x*y
  # modify input
  input <- reduce2(list('\\+', '\\*'), list('/', '-'),
                   .init = input,
                   function(.x, .y, .z) map(.x, function(.a) str_replace_all(.a, .y, .z )))
  
  # final answer
  map_dbl(input, ~ eval(parse(text = .x))) %>% 
    sum()
}

print(answer18_2(input), digits = 14)
```

### Day 19
### [](https://adventofcode.com/2020/day/19)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


### Day 20
### [](https://adventofcode.com/2020/day/20)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


### Day 21
### [--- Day 21: Allergen Assessment ---](https://adventofcode.com/2020/day/21)

#### Part-1
```{r}
input <- read_lines('input21.txt')

answer21_1 <- function(input){
  # patterns
  patterns <- '{ingr=[^\\(]*} (contains {aller=[^\\)]*})'
  
  # read data
  dat <- unglue_data(input, patterns=patterns)
  # all allergans
  all_allregans <- str_split(dat$aller, ', ') %>% unlist %>% unique %>% sort()
  
  # modify data
  dat <- dat %>% 
    mutate(ingr = str_split(ingr, ' '))
  
  # build a list
  map_list <- map(all_allregans, ~ str_detect(dat$aller, paste0('\\b', .x, '\\b')))
  
  
  alle_code <- map(map_list, ~ reduce(dat$ingr[.x], intersect)) %>% 
    unlist %>% unique
  
  {dat$ingr %>% unlist %>% length()} - {map_int(alle_code, ~ sum(dat$ingr %>% unlist == .x)) %>% sum()}
}

answer21_1(input)
```

#### Part-2
```{r}
answer21_2 <- function(input){
  # patterns
  patterns <- '{ingr=[^\\(]*} (contains {aller=[^\\)]*})'
  # data
  dat <- unglue_data(input, patterns=patterns)
  # all allergans
  all_allregans <- str_split(dat$aller, ', ') %>% unlist %>% unique %>% sort()
  # modify data
  dat <- dat %>% 
    mutate(ingr = str_split(ingr, ' '))
  # intermediate list-1
  map_list <- map(all_allregans, ~ str_detect(dat$aller, paste0('\\b', .x, '\\b')))
  # intermediate list2
  list2 <- map(map_list, ~ reduce(dat$ingr[.x], intersect))
  names(list2) <- all_allregans
  list2 <- list2[order(map_int(list2, length))]
  
  # final list for answer
  my_list <- c()
  
  # while loop
  while(1 %in% map_int(list2, length)){
    my_list <- append(my_list, unlist(list2[map_int(list2, length) ==1]))
    list2 <- map(list2, ~ .x[!.x %in% my_list])
  }
  # answer
  paste0(my_list[order(names(my_list))], collapse = ',')
}

answer21_2(input)
```


### Day 22
### [--- Day 22: Crab Combat ---](https://adventofcode.com/2020/day/22)

#### Part-1
```{r}
input <- scan("input22.txt", "character", sep = "\n")

answer22_1 <- function(input){
  # read file correctly
  players <- grep("^Player", input)
  
  player1 <- input[(players[1]+1):(players[2]-1)] %>% as.integer()
  player2 <- input[(players[2]+1):(length(input))] %>%  as.integer()
  
  # while loop
  while(all(map_int(list(player1, player2), length) > 0)){
    x <- player1[1]
    player1 <- player1[-1]
    y <- player2[1]
    player2 <- player2[-1]
    if(x > y){
      player1 <- append(player1, c(x, y))
    } else {
      player2 <- append(player2, c(y, x))
    }
  }
  # answer
  sum(c(player1, player2) * 50:1)
}

answer22_1(input)
```

#### Part-2
```{r}

```


### Day 23
### [](https://adventofcode.com/2020/day/23)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


### Day 24
### [](https://adventofcode.com/2020/day/24)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


### Day 25
### [](https://adventofcode.com/2020/day/25)

#### Part-1
```{r}

```

#### Part-2
```{r}

```


